---
title: "DenverSugar Design"
author: "B. Saul"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DenverSugar Design}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

> NOTE: The terms "vector" and "variable" are (mostly) used interchangably in this document.

# Design Goals

`DenverSugar` provides an extensible set of data types that in themselves extend certain `R` vector classes to be useful in NoviSci's suite of analytic applications by providing vectors with:

   * data types that align with common statistical data types (e.g. "binary", "continous", "count", "time to event", etc);
   * a study design `context` with information about how the variable relates to a study design;
   * a data `description` with relevant summary statistics of the variable;
   * methods that update the `description` when a vector is subset or modified in certain ways;
   * various useful utilities such as predicate functions for identifying variables based on the data type or `context`.
   
## Implementation

The package relies heavily on the [`vctrs`](https://vctrs.r-lib.org/index.html) package whose goals are:

* To propose `vec_size()` and `vec_type()` as alternatives to `length()` and `class()`; `vignette("type-size")`. These definitions are paired with a framework for type-coercion and size-recycling.
* To define type- and size-stability as desirable function properties, use them to analyse existing base function, and to propose better alternatives; `vignette("stability")`. This work has been particularly motivated by thinking about the ideal properties of `c()`, `ifelse()`, and `rbind()`.
* To provide a new `vctr` base class that makes it easy to create new S3 vectors; `vignette("s3-vector")`. vctrs provides methods for many base generics in terms of a few new vctrs generics, making implementation considerably simpler and more robust.

## A quick example

Each data type provided by `DenverSugar` (described in more detail below) have constructor functions that begin with `v_<type>`. For example, `v_binary` creates binary ($\{0, 1\}$) data from R's `logical` type.

```{r}
library(DenverSugar)

x <- v_binary(c(TRUE, TRUE, TRUE, FALSE))

str(x)
```

The `v_binary` data type prints `0`s and `1`s but the underlying data is `logical`:

```{r}
x
vctrs::vec_data(x)
```

The data type includes some useful utilities such as prettying certain parts of the `description` (here the proportion) and a predicate function.

```{r}
is_binary(x)
```

Certain math operations work and pull directly from the `description` where appropriate (rather than recomputing). Note these operations are still under development and should be used with caution:
```{r}
mean(x)
sum(x)

sum(x, x) # See? very experimental

```

Other math/arithmetic operations don't work:
```{r, error=TRUE}
# What do you mean you want to add binary and integer?
x + 2L 

# R's base types are not so safe
vctrs::vec_data(x) + 2L
```

Logical operators work as one might expect: 

```{r}
!x
all(x)
any(x)
```

Here's where the real magic is. 

```{r, error=TRUE}
# vectors can be combined and ...
# subsetting maintains and updates attributes
c(x, !x[1:3])

# But ...
c(x, v_binary(context = context(purpose = "other")))

```

# Data types

The following table describes the proposed data types (not all of these may be available at this time).

| `v_<type>`               | prototype    | support            | 
|--------------------------|--------------|--------------------|
| `v_binary`               | `logical`    | $\{0, 1\}$         |
| `v_count `               | `integer`    | $(0, 1, 2, \dots)$ |
| `v_continuous`           | `double`     | $\mathcal{R}$      |
| --`v_continuous_nonneg`  | `double`     | $\mathcal{R}^+$    |
| --`v_event_time`         | `double`     | $\mathcal{R}^+$    |
| `v_categorical_nom`      | `factor`     | |
| `v_categorical_ord`      | `ordered`    | |
| `v_date`                 | `Date`       | |
| `v_character`            | `character`  | |

# Usage in `tibble`

```{r}
library(dplyr)
n <- 100 

dt <- tibble::tibble(
  y1 = v_binary(as.logical(rbinom(n, 1, 0.25)), context = context(purpose = "outcome")),
  y2 = v_event_time(runif(n, 1, 100), context = context(purpose = "outcome")),
  y3 = v_continuous(rnorm(n), context = context(purpose = "outcome")),
  
  
  !!! purrr::map(
    .x  = 1:10, 
    .f = ~ v_binary(as.logical(rbinom(n, 1, 0.25)), 
                    context = context(purpose = "covariate"))) %>%
    setNames(paste0("x", 1:10))
)

dt
```

Selecting columns based on data type:

```{r}
dt %>%
  select_if(is_binary)
```

Selecting columns based on context:

```{r}
dt %>%
  select_if(is_outcome)
```

Obtain summary statistics for all outcomes in each subgroups 

```{r}
dt %>%
 group_split(x1) %>%
 purrr::map(.x = ., ~ .x %>% select_if(is_outcome) ) %>%
 purrr::map(describe)
```
