---
title: "Using stype lenses"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using stype lenses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# `lenses` and `stype`

The `stype` package provides [`lenses`](https://cfhammill.github.io/lenses/) for getting and settings two types of data:

1. the attributes of an individual `stype` vector (e.g. a `context`, `purpose` or `internal_name`)
2. sets of `stype` vectors within a `list`-like structure (such as a `data.frame` or `tibble`)

## `view` and `set` elements of a `stype` vector

```{r, message=FALSE, warning=FALSE}
library(stype)
library(dplyr)
```

A `lens` consists of two functions: `view` and `set`. `set` can be used to update parts of a `stype` vector.

```{r}
x <-  v_binary() %>%
  set(derivation_l, "some derivation") %>% 
  set(long_label_l, "a new label") %>% 
  set(study_role_l, "outcome")
x
```

And the `view` can be used to get parts of a vector:
```{r}
view(x, derivation_l)
view(x, long_label_l)
view(x, study_role_l)
```

In most cases, `stype` also has `get_*` functions that mimic `view`:

```{r}
get_derivation(x)
get_long_label(x)
```

## `view` and `set` `stype` vectors within a list-like structure

Lenses for `list`-like structures can be used to `view` and `set` vectors within a data.frame. In this example, we set all `continuous` variables to be covariates, `binary` to exposures, and `count` to outcome:

```{r}
df <- tibble(
  x1 = v_binary(),
  x2 = v_continuous(),
  x3 = v_continuous(),
  x4 = v_count()
)

df <- df %>%
  over_map(continuous_l, function(x) set(x, study_role_l, "covariate"))%>%
  over_map(binary_l, function(x) set(x, study_role_l, "exposure")) %>%
  over_map(count_l, function(x) set(x, study_role_l, "outcome"))
 
view(df$x1, study_role_l)
view(df$x2, study_role_l)
view(df$x3, study_role_l) 
view(df$x4, study_role_l) 
```
