---
title: "Workflow for weighting stypes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow for weighting stypes}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(stype)
library(dplyr)
library(lenses)
```

# `weight`ing a `stype` to update the `data_summary`

The `weight` function updates a the `data_summary` of a `stype` vector with a weight.

## a single `stype` vector

```{r}
x <- v_binary(c(TRUE, TRUE, TRUE, FALSE))
w <- c(10, 10, 1, 1)
get_data_summary(x)

z <- weight(x, w)
class(w)
get_data_summary(z)
```

However, the weights are not part of the data, thus when subsetting the vector the `data_summary` no longer includes the weighted summaries:

```{r}
get_data_summary(z[1:2])
```

## a list of `stype` vectors

Here, we use `lenses` to `weight` only certain variables within a `tibble`.


```{r}
weight_var_l <- weight_l %.% lenses::index_l(1) %.% vec_data_l

otcm <- context(purpose = purpose(study_role = "outcome"))
# note the different ways of setting the context
dt <- tibble::tibble(
  x1 = v_binary(c(TRUE, FALSE, TRUE), context  = otcm),
  x2 = v_binary(c(FALSE, TRUE, FALSE)) %>% set(context_l, otcm),
  x3 = v_continuous(c(1.5, 1.6, 0.1), context = otcm),
  x4 = v_continuous(c(1.5, 1.6, 0.1)) %>%
    set(context_l, context(purpose = purpose(study_role = "weight"))),
)

tail(view(dt[[1]], data_summary_l), 2)

# Just weight binary variables
dt1 <- over_map(dt, outcome_l %.% binary_l, function(x){ 
  weight(x, view(dt, weight_var_l))
})
dt1
tail(view(dt1[[1]], data_summary_l), 2)
tail(view(dt1[[3]], data_summary_l), 2)

# Weight all outcomes
dt2 <- over_map(dt, outcome_l, function(x){ 
  weight(x, view(dt, weight_var_l))
})

tail(view(dt2[[1]], data_summary_l), 2)
tail(view(dt2[[3]], data_summary_l), 2)
```

# Estimating cumulative risk

```{r}
ctimes <- list(
   v_event_time(c(5, 6, 10, NA_integer_, 1, NA_integer_, 19), 
                internal_name = "cA"),
   v_event_time(c(4, 1, 15, NA_integer_, NA_integer_, NA_integer_, 21), 
                internal_name = "cB")
)

otimes <- list(
  v_event_time(c(2, 6, 11, 12, NA_integer_, NA_integer_, 25),
               internal_name = "oA"),
  v_event_time(c(1, NA_integer_, 10, NA_integer_, NA_integer_, NA_integer_, 23), 
               internal_name = "oB")
)

x1 <- v_rcensored(
  outcomes = otimes, censors = ctimes, end_time = 15,
  extra_descriptors = list(
    crisk = function(x) { 
      list( riskimator::cumrisk(x, w = riskimator::product_limit) )
    } ))


get_data_summary(x1, "crisk")
get_data_summary(x1[c(3,5,6)], "crisk")
```

## Cumrisk in a data pipeline

```{r, eval= FALSE}
library(riskimator)

dt <- tibble(
  x1 = v_rcensored(outcomes = otimes, censors = ctimes, end_time = 15),
  x2 = v_rcensored(outcomes = otimes[[1]], end_time = 15),
  x3 = v_binary(as.logical(rbinom(7, 1, prob = 0.5)),
                context  = context(purpose = purpose(study_role = "covariate"))),
)

est_cph <- function(x, df){
  cph <- survival::coxph(
    formula = as_Surv(x, censor_as_event = TRUE) ~ as_canonical(x3),
    data = df)
  exp(-predict(cph, type = "expected"))
}
# est_cph(x = dt$x1, df = dt)

weight_rcensored_outcomes <- function(dt){
  over_map(dt, rcensored_l, function(x) {
    set(x, data_summary_l, 
        c(view(x, data_summary_l), 
          list(crisk = cumrisk(x, w = est_cph, df = dt))))
  })
}
# 
dt2 <- weight_rcensored_outcomes(dt)
view(dt2$x1, data_summary_l)

```


